---
title: "Homework 4 - Insurance"
author: "Elina Azrilyan"
date: "April 22, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(MASS)
library(pROC)
```

Reading in cleaned up output dataset:

```{r}
ins_df <- read.csv( "https://raw.githubusercontent.com/charlsjoseph/Data621/master/Data621-Assignment4/insurance_tf_train.csv")[-1]
test_set <- read.csv("https://raw.githubusercontent.com/charlsjoseph/Data621/master/Data621-Assignment4/insurance_tf_test.csv")[-1]
df_eval <- read.csv( "https://raw.githubusercontent.com/charlsjoseph/Data621/master/Data621-Assignment4/insurance_tf_eval.csv")[-1]
```

###Multiple Linear Regression

####Model 1

We will try to use all the variables for our initial regression model, to identify which appear to show significance.

```{r}
ins_df$TARGET_FLAG <- NULL
reg1<-lm(formula = TARGET_AMT ~ ., data = ins_df)
summary(reg1)
```

There are a lot of variable with low signifiance and high p-value, so we will try to only leave the variable with high significance for the next model.

```{r}
reg1_1<-lm(formula = TARGET_AMT ~ KIDSDRIV + PARENT1.Yes + MSTATUS.Yes + TRAVTIME + CAR_USE.Private + TIF + CAR_TYPE.Minivan + REVOKED.Yes + MVR_PTS + CAR_AGE + URBANICITY.Highly.Urban..Urban, data = ins_df)
summary(reg1_1)
```

We can see improvements in R-Squared value after we narrowed the model to highly significant variables. It is however still rather low. Only 7% of the variation in the data can be explained by this model. 

We will now use this regression model to predict values for the evaluation data set. 
```{r}
reg_pred <- predict(reg1_1, df_eval, type="response",se.fit=FALSE)
# se.fit=FALSE
df_eval$TARGET_AMT <- reg_pred

write.csv(df_eval,"eval_results.csv", row.names = FALSE)
```